\emph{Ab initio} Calculations of Intramolecular Exciton Transfer with Reduced Modes in Donor-Bridge-Acceptor Species

### 2016-04-25 Chapters/chap4/chap4.tex
Defense passed!

\begin{figure*}[!h]
\includegraphics[width=0.45\columnwidth]{Chapters/chap4/Images/parabola.jpg}
\caption{Calculated energies of the CT (black), CSS (blue) and 3NAP (red) states along the NAP side C$\equiv$C coordinate in the ground state geometries of the molecules studied, in CH\textsubscript{2}Cl\textsubscript{2}. a, PTZ-CH\textsubscript{2}-Pt-NAP. b, PTZ-Pt-NAP. c, OMe-PTZ-Pt-NAP. The first and second vibrational energy levels for the asymmetric C$\equiv$C stretch are shown for the CT state. The energy gaps from $\nu_0$ to the intersection between CT and CSS are shown in the insets. There is a clear trend concerning the energy gap: PTZ-CH\textsubscript{2}-Pt-NAP < PTZ-Pt-NAP < OMe-PTZ-Pt-NAP, reflecting the trend experimentally observed in the efficiency of the infrared control effect. Reprinted with permission from Ref.~\cite{delor2015mechanism}.\label{PTZparabolas}} \end{figure*}

@@

\begin{figure}[!t]
\includegraphics[width=\columnwidth]{Chapters/chap4/Images/energy_diagram.pdf}
\caption{The energy diagram of triplet states at \textsuperscript{3}NAP and CT state geometries. The electron/hole distribution are shown as well (green for electron, blue for hole). The arrows indicate the transitions we calculate.
}\label{ene_diagram}
\end{figure}

@@

\begin{figure*}[!ht]
\includegraphics[width=\columnwidth]{Chapters/chap4/Images/kTCLME-VS-EXPT.pdf}
\caption{Comparisons between theoretical rate constants and the experimental rates. Experimental rates are calculated as yield of product/lifetime of reactant, using values from ~Ref. \cite{delor2015mechanism}.  The ``Marcus'' rates are calculated using Marcus theory with the diabatic couplings at according geometries. The ``Marcus (mean V)'' rates are calculated using mean diabatic couplings along the interpolated coordinate in Fig. \ref{interCondon}. Similarly, the ``TCLME'' and ``TCLME (mean V)'' rates are calculated using the TCLME approach, with all normal modes used. The ``PLM'' rates are calculated using only the primary mode. Our approach performs better than Marcus theory. The CT geometry gives good ``exact'' results, its primary mode rates also agree well. While for the \textsuperscript{3}NAP geometry, using mean value of diabatic couplings along the reaction coordinate significantly improves the performance.\label{rateComp}}
\end{figure*}

\begin{figure}[!t]
\subfloat[]{\includegraphics[width=0.5\columnwidth]{Chapters/chap4/Images/corrT12.pdf}}
\subfloat[]{\includegraphics[width=0.5\columnwidth]{Chapters/chap4/Images/corrT13.pdf}}\\
\subfloat[]{\includegraphics[width=0.5\columnwidth]{Chapters/chap4/Images/corrT31.pdf}}
\subfloat[]{\includegraphics[width=0.5\columnwidth]{Chapters/chap4/Images/corrT32.pdf}}
\caption{
Correlation functions of various numbers of projected modes, compared to exact correlation, for (a) CSS $\rightarrow$ \textsuperscript{3}NAP at \textsuperscript{3}NAP geometry, (b) CT $\rightarrow$ \textsuperscript{3}NAP at \textsuperscript{3}NAP geometry, (c) CT $\rightarrow$ \textsuperscript{3}NAP at CT geometry, and (d) CT $\rightarrow$ CSS at CT geometry.
}\label{corrT1T3}
\end{figure}
\subsection{Primary Mode}

@@

\begin{figure}[!t]
\subfloat[]{\includegraphics[width=0.88\columnwidth]{Chapters/chap4/Images/plmT12.png}}\\
\subfloat[]{\includegraphics[width=0.88\columnwidth]{Chapters/chap4/Images/plmT13.png}}
% \subfloat[]{\includegraphics[width=0.88\columnwidth]{Chapters/chap4/Images/plmT31.png}}\\
% \subfloat[]{\includegraphics[width=0.88\columnwidth]{Chapters/chap4/Images/plmT32.png}}
\caption{
Projection of primary mode of (a) CSS $\rightarrow$ \textsuperscript{3}NAP, (b) CT $\rightarrow$ \textsuperscript{3}NAP calculated at \textsuperscript{3}NAP geometry onto the normal modes of \textsuperscript{3}NAP. Embedded molecule shows the atomic displacement vectors of primary mode. If we focus on the C$\equiv$C bonds, PLMs are like
PTZ-Ph-C$\equiv$C-Pt-C$\protect\overset{\longleftrightarrow}{\equiv}$C-NAP
in (a) and
PTZ-Ph-$\protect\overset{\longleftarrow}{C}\equiv\protect\overset{\longrightarrow}{C}$-Pt-$\protect\overset{\longrightarrow}{C}\equiv\protect\overset{\longleftarrow}{C}$-NAP
in (b).}\label{projT1}
\end{figure}

\begin{figure}[!h]
% \subfloat[]{\includegraphics[width=0.89\columnwidth]{Chapters/chap4/Images/plmT12.png}}\\
% \subfloat[]{\includegraphics[width=0.89\columnwidth]{Chapters/chap4/Images/plmT13.png}}
\subfloat[]{\includegraphics[width=0.89\columnwidth]{Chapters/chap4/Images/plmT31.png}}\\
\subfloat[]{\includegraphics[width=0.89\columnwidth]{Chapters/chap4/Images/plmT32.png}}
\caption{
Projection of primary mode of (a) CT $\rightarrow$ \textsuperscript{3}NAP, and (b) CT $\rightarrow$ CSS calculated at CT geometry onto the normal modes of CT. Embedded molecule shows the atomic displacement vectors of primary mode. If we focus on the C$\equiv$C bonds, PLMs are like
PTZ-Ph-$\protect\overset{\longleftarrow}{C}\equiv\protect\overset{\longrightarrow}{C}$-Pt-$\protect\overset{\rightarrow}{C}\equiv\protect\overset{\leftarrow}{C}$-NAP
in (a) and
PTZ-Ph-$\protect\overset{\longleftarrow}{C}\equiv\protect\overset{\longrightarrow}{C}$-Pt-$\protect\overset{\leftarrow}{C}\equiv\protect\overset{\rightarrow}{C}$-NAP
in (b).
}\label{projT3}
\end{figure}

\begin{figure}[!h]
\subfloat{\includegraphics[height=0.8\textheight,keepaspectratio]{Chapters/chap4/Images/t1_174.jpg}}\\
\subfloat{\includegraphics[height=0.8\textheight,keepaspectratio]{Chapters/chap4/Images/t1_175.jpg}}\\
\subfloat{\includegraphics[height=0.8\textheight,keepaspectratio]{Chapters/chap4/Images/t3_174.jpg}}\\
\subfloat{\includegraphics[height=0.8\textheight,keepaspectratio]{Chapters/chap4/Images/t3_175.jpg}}
\caption{From top to bottom: the displacement vectors of the 174th, 175th mode of \textsuperscript{3}NAP state, the 174th, 175th mode of CT state. For clarity, bonds are not shown. They are all the stretching modes of single C$\equiv$C bond, look like
PTZ-Ph-C$\equiv$C-Pt-C$\protect\overset{\longleftrightarrow}{\equiv}$C-NAP in (a) and (c),
PTZ-Ph-C$\protect\overset{\longleftrightarrow}{\equiv}$C-Pt-C$\equiv$C-NAP in (b) and (d).
}\label{NMT1T3}
\end{figure}

As experiments show that the excitation of C$\equiv$C bond radically affected the ET process, we expect the triple bond vibration to dominate the primary modes, at least for the CT state based calculations, since they achieve good agreement. Fig. \ref{projT1} and \ref{projT3} show the projection of primary modes onto the normal modes of corresponding state, {\em i.~e.}, the primary modes calculated at CT geometry are projected onto the normal modes of the CT state, and those at \textsuperscript{3}NAP are projected onto \textsuperscript{3}NAP's normal modes. It turns out that all four primary modes involve obvious C$\equiv$C displacement, and it is verified by the projections. The projections of primary modes are dominated by 1 or 2 normal modes. In Fig. \ref{projT1}(a), it is the 174th normal mode (ordered by increasing frequency) that dominates; in Fig. \ref{projT1}(b), the 174th and 175th modes dominate. Similarly, 174th and 175th modes dominate in Fig. \ref{projT3}(a) and  175th mode in Fig. \ref{projT3}(b). All these normal modes are C$\equiv$C stretching modes, as shown in Fig. \ref{NMT1T3}. We suggest that the primary mode of  CT $\rightarrow$ CSS transition (Fig. \ref{projT3}(b)) is the one involves C$\equiv$C vibration most. It has a component which has the largest weight, > 0.25, in all four primary modes. And that component clearly dominates other normal modes. As for PTZ-CH\textsubscript{2}-Pt-NAP, the CT $\rightarrow$ CSS step can be fully switched off by vibronic excitation, the step is supposed to be the one influenced by C$\equiv$C stretching most. Although PTZ-Pt-NAP does not show such radical change, we expect the the CT $\rightarrow$ CSS to be the most C$\equiv$C involved, as two molecules only differ by a -CH\textsubscript{2}-

@@

\begin{figure*}[!h]
\includegraphics[width=\columnwidth]{Chapters/chap4/Images/interpolation.pdf}
\caption{Diabatic coupling values along the interpolation from the \textsuperscript{3}NAP $\rightarrow$ CT geometries. State 0 and 10 stand for the \textsuperscript{3}NAP and the CT state, respectively.\label{interCondon}}
\end{figure*}

>>>>

\begin{figure*}[]
\includegraphics[width=0.45\columnwidth]{Chapters/chap4/Images/parabola.jpg}
\caption{Calculated energies of the CT (black), CSS (blue) and \textsuperscript{3}NAP (red) states along the NAP side C$\equiv$C coordinate in the ground state geometries of the molecules studied, in CH\textsubscript{2}Cl\textsubscript{2}. a, PTZ-CH\textsubscript{2}-Pt-NAP. b, PTZ-Pt-NAP. c, OMe-PTZ-Pt-NAP. The first and second vibrational energy levels for the asymmetric C$\equiv$C stretch are shown for the CT state. The energy gaps from $\nu_0$ to the intersection between CT and CSS are shown in the insets. There is a clear trend concerning the energy gap: PTZ-CH\textsubscript{2}-Pt-NAP < PTZ-Pt-NAP < OMe-PTZ-Pt-NAP, reflecting the trend experimentally observed in the efficiency of the infrared control effect. Reprinted with permission from Ref.~\cite{delor2015mechanism}.\label{PTZparabolas}} \end{figure*}

@@

\begin{figure}[]
\includegraphics[width=\columnwidth]{Chapters/chap4/Images/energy_diagram.pdf}
\caption{The energy diagram of triplet states at \textsuperscript{3}NAP and CT state geometries. The electron/hole distribution are shown as well (green for electron, blue for hole). The arrows indicate the transitions we calculate.
}\label{ene_diagram}
\end{figure}

@@

\begin{figure*}[]
\includegraphics[width=\columnwidth]{Chapters/chap4/Images/kTCLME-VS-EXPT.pdf}
\caption{Comparisons between theoretical rate constants and the experimental rates. Experimental rates are calculated as yield of product/lifetime of reactant, using values from ~Ref. \cite{delor2015mechanism}.  The ``Marcus'' rates are calculated using Marcus theory with the diabatic couplings at according geometries. The ``Marcus (mean V)'' rates are calculated using mean diabatic couplings along the interpolated coordinate in Fig. \ref{interCondon}. Similarly, the ``TCLME'' and ``TCLME (mean V)'' rates are calculated using the TCLME approach, with all normal modes used. The ``PLM'' rates are calculated using only the primary mode. Our approach performs better than Marcus theory. The CT geometry gives good ``exact'' results, its primary mode rates also agree well. While for the \textsuperscript{3}NAP geometry, using mean value of diabatic couplings along the reaction coordinate significantly improves the performance.\label{rateComp}}
\end{figure*}

\begin{figure}[]
\subfloat[]{\includegraphics[width=0.5\columnwidth]{Chapters/chap4/Images/corrT12.pdf}}
\subfloat[]{\includegraphics[width=0.5\columnwidth]{Chapters/chap4/Images/corrT13.pdf}}\\
\subfloat[]{\includegraphics[width=0.5\columnwidth]{Chapters/chap4/Images/corrT31.pdf}}
\subfloat[]{\includegraphics[width=0.5\columnwidth]{Chapters/chap4/Images/corrT32.pdf}}
\caption{
Correlation functions of various numbers of projected modes, compared to exact correlation, for (a) CSS $\rightarrow$ \textsuperscript{3}NAP at \textsuperscript{3}NAP geometry, (b) CT $\rightarrow$ \textsuperscript{3}NAP at \textsuperscript{3}NAP geometry, (c) CT $\rightarrow$ \textsuperscript{3}NAP at CT geometry, and (d) CT $\rightarrow$ CSS at CT geometry.
}\label{corrT1T3}
\end{figure}

\subsection{Primary Mode}

@@

\begin{figure}[]
\subfloat[]{\includegraphics[width=0.88\columnwidth]{Chapters/chap4/Images/plmT12.png}}\\
\subfloat[]{\includegraphics[width=0.88\columnwidth]{Chapters/chap4/Images/plmT13.png}}
% \subfloat[]{\includegraphics[width=0.88\columnwidth]{Chapters/chap4/Images/plmT31.png}}\\
% \subfloat[]{\includegraphics[width=0.88\columnwidth]{Chapters/chap4/Images/plmT32.png}}
\caption{
Projection of primary mode of (a) CSS $\rightarrow$ \textsuperscript{3}NAP, (b) CT $\rightarrow$ \textsuperscript{3}NAP calculated at \textsuperscript{3}NAP geometry onto the normal modes of \textsuperscript{3}NAP. Embedded molecule shows the atomic displacement vectors of primary mode. If we focus on the C$\equiv$C bonds, PLMs are like
PTZ-Ph-C$\equiv$C-Pt-C$\protect\overset{\longleftrightarrow}{\equiv}$C-NAP
in (a) and
PTZ-Ph-$\protect\overset{\longleftarrow}{C}\equiv\protect\overset{\longrightarrow}{C}$-Pt-$\protect\overset{\longrightarrow}{C}\equiv\protect\overset{\longleftarrow}{C}$-NAP
in (b).}\label{projT1}
\end{figure}

\begin{figure}[]
% \subfloat[]{\includegraphics[width=0.89\columnwidth]{Chapters/chap4/Images/plmT12.png}}\\
% \subfloat[]{\includegraphics[width=0.89\columnwidth]{Chapters/chap4/Images/plmT13.png}}
\subfloat[]{\includegraphics[width=0.89\columnwidth]{Chapters/chap4/Images/plmT31.png}}\\
\subfloat[]{\includegraphics[width=0.89\columnwidth]{Chapters/chap4/Images/plmT32.png}}
\caption{
Projection of primary mode of (a) CT $\rightarrow$ \textsuperscript{3}NAP, and (b) CT $\rightarrow$ CSS calculated at CT geometry onto the normal modes of CT. Embedded molecule shows the atomic displacement vectors of primary mode. If we focus on the C$\equiv$C bonds, PLMs are like
PTZ-Ph-$\protect\overset{\longleftarrow}{C}\equiv\protect\overset{\longrightarrow}{C}$-Pt-$\protect\overset{\rightarrow}{C}\equiv\protect\overset{\leftarrow}{C}$-NAP
in (a) and
PTZ-Ph-$\protect\overset{\longleftarrow}{C}\equiv\protect\overset{\longrightarrow}{C}$-Pt-$\protect\overset{\leftarrow}{C}\equiv\protect\overset{\rightarrow}{C}$-NAP
in (b).
}\label{projT3}
\end{figure}

\begin{figure}[]
\subfloat{\includegraphics[width=0.85\columnwidth]{Chapters/chap4/Images/t1_174.png}}\\
\subfloat{\includegraphics[width=0.85\columnwidth]{Chapters/chap4/Images/t1_175.png}}\\
\subfloat{\includegraphics[width=0.85\columnwidth]{Chapters/chap4/Images/t3_174.png}}\\
\subfloat{\includegraphics[width=0.85\columnwidth]{Chapters/chap4/Images/t3_175.png}}
\caption{From top to bottom: the displacement vectors of the 174th, 175th mode of \textsuperscript{3}NAP state, the 174th, 175th mode of CT state. For clarity, bonds are not shown. They are all the stretching modes of single C$\equiv$C bond, look like
PTZ-Ph-C$\equiv$C-Pt-C$\protect\overset{\longleftrightarrow}{\equiv}$C-NAP in (a),
PTZ-Ph-C$\protect\overset{\longleftrightarrow}{\equiv}$C-Pt-C$\equiv$C-NAP in (b),
PTZ-Ph-$\protect\overset{\leftarrow}{C}\equiv\protect\overset{\rightarrow}{C}$-Pt-$\protect\overset{\longrightarrow}{C}\equiv\protect\overset{\longleftarrow}{C}$-NAP
 in (c), and
PTZ-Ph-$\protect\overset{\longleftarrow}{C}\equiv\protect\overset{\longrightarrow}{C}$-Pt-$\protect\overset{\leftarrow}{C}\equiv\protect\overset{\rightarrow}{C}$-NAP
 in (d).
}\label{NMT1T3}
\end{figure}

As experiments show that the excitation of C$\equiv$C bond radically affected the ET process, we expect the triple bond vibration to dominate the primary modes, at least for the CT state based calculations, since they achieve good agreement. Fig. \ref{projT1} and \ref{projT3} show the projection of primary modes onto the normal modes of corresponding state, {\em i.~e.}, the primary modes calculated at CT geometry are projected onto the normal modes of the CT state, and those at \textsuperscript{3}NAP are projected onto \textsuperscript{3}NAP's normal modes. It turns out that all four primary modes involve obvious C$\equiv$C displacement, and it is verified by the projections. The projections of primary modes are dominated by 1 or 2 normal modes. In Fig. \ref{projT1}(a), it is the 174th normal mode (ordered by increasing frequency) that dominates; in Fig. \ref{projT1}(b), the 174th and 175th modes dominate. Similarly, 174th and 175th modes dominate in Fig. \ref{projT3}(a) and  175th mode in Fig. \ref{projT3}(b).

All these normal modes are C$\equiv$C stretching modes, as shown in Fig. \ref{NMT1T3}. We suggest that the primary mode of  CT $\rightarrow$ CSS transition (Fig. \ref{projT3}(b)) is the one involves C$\equiv$C vibration most. It has a component which has the largest weight, > 0.25, in all four primary modes. And that component clearly dominates other normal modes. Experimentally, after the IR pump is applied, the CT $\rightarrow$  \textsuperscript{3}NAP transition is relatively accelerated compared to the CT $\rightarrow$ CSS transition. By comparing the primary modes of the two transitions at CT geometry, we see that the primary mode of CT $\rightarrow$  \textsuperscript{3}NAP transition  is mainly antisymmetric stretching of two C$\equiv$C bonds, while that of CT $\rightarrow$ CSS transition is symmetric. As a result, the primary mode of CT $\rightarrow$  \textsuperscript{3}NAP is more IR active than that of CT $\rightarrow$ CSS transition. It may explain why CT $\rightarrow$  \textsuperscript{3}NAP  transition is relatively accelerated with IR pump.
% As for PTZ-CH\textsubscript{2}-Pt-NAP, the CT $\rightarrow$ CSS step can be fully switched off by vibronic excitation, the step is supposed to be the one influenced by C$\equiv$C stretching most. Although PTZ-Pt-NAP does not show such radical change, we expect the the CT $\rightarrow$ CSS to be the most C$\equiv$C involved, as two molecules only differ by a -CH\textsubscript{2}-

@@

\begin{figure*}[t]
\includegraphics[width=\columnwidth]{Chapters/chap4/Images/interpolation.pdf}
\caption{Diabatic coupling values along the interpolation from the \textsuperscript{3}NAP $\rightarrow$ CT geometries. State 0 and 10 stand for the \textsuperscript{3}NAP and the CT state, respectively.\label{interCondon}}
\end{figure*}
